Adaptive group-regularized logistic elastic net regression Supplementary material

### 2019-02-05 docs/supplement_biostatistics.Rnw
adjusted alzheimer data analysis

\documentclass[a4paper,hidelinks]{article}
\usepackage{bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,bbm,algorithm,algpseudocode,pgfplotstable,threeparttable, booktabs,mathtools,dsfont,parskip}

@@

% syntax and readability
\renewcommand{\(}{\left(}
\renewcommand{\)}{\right)}
\renewcommand{\[}{\left[}
\renewcommand{\]}{\right]}
\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}} 
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

% settings
\graphicspath{{../figs/}}
\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
\setlength{\textwidth}{15cm}
\setlength{\parindent}{0pt}
\title{Adaptive group-regularized logistic elastic net regression Supplementary material}
\date{}
\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: \href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Carel F.W. Peeters$^{1}$, Aad W. van der Vaart$^{2}$ \\ and Mark A. van de Wiel$^{1,3}$}

\begin{document}
\maketitle
\noindent
1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health research institute, VU University Medical Center, PO Box 7057, 1007 MB
Amsterdam, The Netherlands\\
2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\
3. Department of Mathematics, VU University, Amsterdam, The Netherlands \\

\textbf{Keywords}: Empirical Bayes; High-dimensional data; Prediction; Variational Bayes

\textbf{Software available from}: \url{https://CRAN.R-project.org/package=gren}

\section{Introduction}
This document contains supplementary material to the paper `Adaptive group-regularized logistic elastic net regression'. Section \ref{sec:enetprior} contains extra details on the elastic net prior distribution. Section \ref{sec:objectivefunction} gives the objective function that is maximised in the variational Bayes procedure introduced in Section 3 of the Main Document. Section \ref{sec:polyagamma} gives details on the P{\'o}lya-Gamma parametrisation introduced in Section 3 of the Main Document. The derivations of the variational posterior used in the Main Document are shown in Section \ref{sec:variationaldistributions}. Section \ref{sec:computationalcomplexity} shows how we may reduce the computational complexity of our variational Bayes procedure. Section \ref{sec:expectedloglikelihood} elaborates on the Variational Bayes expected joint log likelihood that is maximised to estimate the penalty multipliers as explained in Section 3 of the Main Document. The full procedure of the proposed method is shown in Section \ref{sec:fullestimationprocedure}. We apply our method to a study on microRNAs in colorectal cancer data \cite[]{neerincx_combination_2018} in Section \ref{sec:appcolon} . Finally, we introduce the ridge and lasso versions of the proposed method in Section \ref{sec:evbridgelasso}.

\section{The elastic net prior}\label{sec:enetprior}
\subsection{Density function}
The density function of $\beta$, drawn from an elastic net prior distribution, is given by:
$$
f(\beta) = g(\lambda_1,\lambda_2) \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\].
$$
The normalizing constant here is calculated as:
$$
g(\lambda_1,\lambda_2) =\left\{ \int_{-\infty}^{\infty} \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta \right\}^{-1}= \frac{\sqrt{\lambda_2}}{2} \phi \( \frac{\lambda_1}{2 \sqrt{\lambda_2}} \) \Phi \(\frac{-\lambda_1}{2 \sqrt{\lambda_2}} \)^{-1},
$$
with $\phi(x)$ and $\Phi(x)$ the standard normal density and distribution functions of $x$, respectively.

\subsection{Expectation and variance}
The expectation and variance of $\beta$ are:
\begin{align*}
\E[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta \cdot \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta=0, \\
\V[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta^2 \cdot \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta - \E[\beta]^2\\ &= \frac{\lambda_1^2}{4 \lambda_2^2} + \frac{1}{\lambda_2} - \frac{\lambda_1}{2 \lambda_2^{3/2}} \phi \( \frac{\lambda_1}{2 \sqrt{\lambda_2}} \) \Phi \(\frac{-\lambda_1}{2 \sqrt{\lambda_2}} \)^{-1}.
\end{align*}

\subsection{Simulation}
Generating samples from the elastic net prior is not straightforward. \cite{li_bayesian_2010} show that the elastic net prior may be written as a scale mixture of normals, with mixing parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} \tr$: 
\begin{align*}
\beta | \tau & \sim \mathcal{N} \left(0,\frac{1}{\lambda_2} \frac{\tau - 1}{\tau} \right), \\
\tau & \sim \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2}{\lambda_1^2}, \left(1,\infty \right) \right),
\end{align*}
where $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$. So with an efficient truncated gamma sampler available we may sample from the prior using any standard software package that includes a Gaussian sampling routine. More specifically, we require a a sample from the $\mathcal{TG} \left( \frac{1}{2},8 \lambda2/\lambda_1^2, \left(1,\infty \right) \right)$ distribution. The CDF of this distribution is given by:
$$
F_{T}(\tau) =	
\begin{dcases}
0, & \text{if }\tau < 1 \\
1 - \frac{\Phi\(-\frac{\lambda_1}{2} \sqrt{\frac{\tau}{\lambda_2}} \)}{\Phi\(-\frac{\lambda_1}{2 \sqrt{\lambda_2}} \)}, & \text{if } \tau \geq 1.
\end{dcases}
$$

Since $F_{T}(\tau)$ is continuous and strictly monotonically increasing for $\tau \geq 1$, the quantile function may be written as a function of the probability $p \in (0,1)$:
$$
Q_{T}(p) =	F^{-1}_{T}(p) = \frac{4 \lambda_2}{\lambda_1^2} \Phi^{-1} \[ (1 - p) \Phi\(-\frac{\lambda_1}{2 \sqrt{\lambda_2}} \)\]^2,
$$
where $\Phi^{-1}(p)$ is the quantile function of the standard normal distribution, i.e, the probit function. Employing inverse transform sampling for the truncated gamma distribution, an elastic net prior sampling scheme is now:
\begin{algorithmic}
	\State{Generate $U \sim \mathcal{U} (0,1)$}
	\State{Set $\tau = \frac{4 \lambda_2}{\lambda_1^2} \Phi^{-1} \[ (1 - u) \Phi\(-\frac{\lambda_1}{2 \sqrt{\lambda_2}} \)\]^2$}
	\State{Set $\nu^2= \frac{\tau - 1}{\tau \lambda_2}$}
	\State{Generate $\beta \sim \mathcal{N}(0,\nu^2)$}
	\State{\Return{$\beta$}}
\end{algorithmic}

\section{Objective function}\label{sec:objectivefunction}
The lower bound on the marginal likelihood, or evidence lower bound (ELBO), at iteration $t$ is given by:
\begin{subequations}
	\begin{align}
	\text{ELBO} (Q^{(t)}) & = \E_{Q^{(t)}} [\log \L (\y, \bomega, \bbeta, \btau)] - \E_{Q^{(t)}} [\log Q (\bomega, \bbeta, \btau)] \\
	& = \E_{Q^{(t)}} [\log \L (\y ; \bbeta)] + \E_{Q^{(t)}} [\log \pi(\bomega | \bbeta)] - \E_{Q^{(t)}} [\log q(\bomega)] \label{eq:elboomega} \\
	& \,\,\,\,\, + \E_{Q^{(t)}} [\log \pi(\bbeta|\btau)] - \E_{Q^{(t)}} [\log q(\bbeta)] \label{eq:elbobeta}\\
	& \,\,\,\,\, + \E_{Q^{(t)}} [\log \pi(\btau)] - \E_{Q^{(t)}} [\log q(\btau)] \label{eq:elbotau}.
	\end{align}
\end{subequations}

We start with the computation of (\ref{eq:elboomega}). To this end we define:
$$
h(m_i, \omega_i) := \sum_{k=0}^{\infty} (-1)^k \frac{\Gamma(k+m_i}{k + 1}\frac{2k +m_i}{\sqrt{2 \pi \omega_i^3}}\exp\[-\frac{(2k + m_i)^2}{8 \omega_i} \].
$$
Let $\c = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix} \tr$. Furthermore, in the following, all expectations and variances are with respect to $Q^{(t)}$. Next we write out the three terms in (\ref{eq:elboomega}):
\begin{subequations} \label{eq:elboomega2}
	\begin{align}
	\E [\log \L (\y ; \bbeta)] & = \sum_{i=1}^n \log \binom{m_i}{y_i} + \y \tr \X \E [\bbeta] - \m \tr \E \left\{ \log [1 + \exp(\x_i \tr \bbeta)] \right\}, \\
	\E [\log \pi(\bomega | \bbeta)] & = \m \tr \E \left\{ \log [1 + \exp(\x_i \tr \bbeta)] \right\} - n \log 2 - \frac{1}{2} \m \tr \X \E [\bbeta] \\
	& \,\,\,\,\, - \sum_{i=1}^n \log \Gamma (m_i) - \frac{1}{2} \sum_{i=1}^n [ ( \x_i \tr \E [\bbeta] )^2 + \x_i \tr \V [\bbeta] \x_i ] \E [\omega_i] \\
	& \,\,\,\,\,+ \sum_{i=1}^n \E \left\{ \log [ h(m_i, \omega_i) ] \right\}, \\
	\E [\log q(\bomega)] & =  \m \tr \log [1 + \exp(\c)] - n \log 2 - \frac{1}{2} \m \tr \c - \sum_{i=1}^n \log \Gamma (m_i) - \frac{1}{2} \sum_{i=1}^n c_i^2 \E [\omega_i] \\
	& \,\,\,\,\, + \sum_{i=1}^n \E \left\{ \log [ h(m_i, \omega_i) ] \right\},
	\end{align}
\end{subequations}
where all mathematical operations on vectors and matrices are element wise. After combining the terms in (\ref{eq:elboomega2}) and substituting the expectations and variances given in the Main Document, several terms cancel to give:
\begin{subequations} \label{eq:elboomega3}
	\begin{align}
	\E_{Q^{(t)}} & [\log \L (\y ; \bbeta)] + \E_{Q^{(t)}} [\log \pi(\bomega | \bbeta)] - \E_{Q^{(t)}} [\log q(\bomega)] \\ & = \sum_{i=1}^n \log \binom{m_i}{y_i} + ( \y - \frac{1}{2} \m) \tr \X \bmu^{(t)} + \m \tr \left\{ \frac{1}{2} \c^{(t)} - \log [1 + \exp(\c^{(t)})] \right\} \label{eq:elboomega4}\\
	& \,\,\,\,\, + \frac{1}{4} \sum_{i=1}^n \frac{m_i}{c_i^{(t)}} \tanh \(\frac{c_i^{(t)}}{2} \)\left\{ (c_i^{(t)})^2 - ( \x_i \tr \bmu^{(t)} )^2 - \x_i \tr \bSigma^{(t)} \x_i \tr \right\}.
	\end{align}
\end{subequations}

Inspection of the updating equations in the Main Document, learns us that the last term in the right-hand side of (\ref{eq:elboomega3}) equals zero, so that we are left with just (\ref{eq:elboomega4}). After a change of variables $\psi_j = \tau_j - 1$, the two terms in (\ref{eq:elbobeta}) are as follows:
\begin{subequations} \label{eq:elbobeta2}
	\begin{align}
	\E [\log \pi (\bbeta | \btau)] & = \frac{1}{2}\sum_{g=1}^G |\mathcal{G}(g)| \log \lambda'_g + \frac{p}{2} \log \lambda_2 - \frac{p}{2} \log (2 \pi) \\
	& \,\,\,\,\, - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j}}\) (\bmu_j^2 + \bSigma_{jj}) \\
	& \,\,\,\,\, - \frac{1}{2} \sum_{j=1}^p \E [\log \psi_j] + \frac{1}{2} \sum_{j=1}^p \E [ \log(\psi_j + 1)], \\
	\E [\log q (\bbeta )] & = -\frac{p}{2} - \frac{p}{2}\log (2 \pi) - \frac{p}{2} \log |\bSigma| .
	\end{align}
\end{subequations}
We combine the terms in (\ref{eq:elbobeta2}) and arrive at the following for (\ref{eq:elbobeta}):
\begin{subequations} \label{eq:elbobeta3}
	\begin{align}
	\E_{Q^{(t)}} & [\log \pi (\bbeta | \btau)] - \E_{Q^{(t)}} [\log q (\bbeta )] \\ 
	& = \frac{1}{2}\sum_{g=1}^G |\mathcal{G}(g)| \log \lambda'^{(t)}_g + \frac{p}{2} \log \lambda_2 + \frac{p}{2} \log |\bSigma^{(t)}| + \frac{p}{2} \\
	& \,\,\,\,\, - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'^{(t)}_g \sum_{j \in \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j^{(t)} } }\) \((\bmu^{(t)}_j)^2 + \bSigma_{jj}^{(t)}\) \\
	& \,\,\,\,\, - \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log \psi_j] + \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [ \log(\psi_j + 1)].
	\end{align}
\end{subequations}

Repeating the same exercise for (\ref{eq:elbotau}) gives:
\begin{subequations} \label{eq:elbotau2}
	\begin{align}
	\E [\log \pi (\btau)] & = p \log \lambda_1 - \frac{p}{2} \log \lambda_2 - \frac{5 p }{2} \log 2 - \frac{p}{2} \log \pi - \frac{p}{2} - \frac{p \lambda_1^2}{8 \lambda_2} \\
	& \,\,\,\,\, - p \log \Phi \( \frac{-\lambda_1}{2 \sqrt{ \lambda_2}} \) - \frac{\lambda_1}{4 \sqrt{\lambda_2}} \sum_{j=1}^p \chi_j^{1/2} - \frac{1}{2} \sum_{j=1}^p \E [\log ( \psi_j + 1)], \\
	\E [\log q (\btau)] & = p \log \lambda_1 - \frac{p}{2} \log \lambda_2 - \frac{3p}{2} \log 2 -\frac{p}{2} \log \pi -\frac{p}{2} - \frac{\lambda_1}{2 \sqrt{\lambda_2}} \sum_{j=1}^p \chi_j^{1/2} \\
	& \,\,\,\,\, + \frac{1}{4} \sum_{j=1}^p \log \chi_j - \frac{1}{2} \sum_{j=1}^p \E [\log \psi_j].
	\end{align}
\end{subequations}
Again, we combine the terms in (\ref{eq:elbotau2}) to get:
\begin{subequations} \label{eq:elbotau3}
	\begin{align}
	\E_{Q^{(t)}} & [\log \pi (\btau)] - \E_{Q^{(t)}} [\log q (\btau)] 
	\\ & = p \log 2 - \frac{p \lambda_1^2}{8 \lambda_2} - p \log \Phi \( \frac{-\lambda_1}{2 \sqrt{ \lambda_2}} \) + \frac{\lambda_1}{4 \sqrt{\lambda_2}} \sum_{j=1}^p \sqrt{\chi_j^{(t)}} - \frac{1}{4} \sum_{j=1}^p \log \chi_j^{(t)} \\
	& \,\,\,\,\, + \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log \psi_j] - \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log ( \psi_j + 1)].
	\end{align}
\end{subequations}

>>>>

\documentclass[a4paper,hidelinks]{article}
\usepackage{bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,bbm,algorithm,algpseudocode,pgfplotstable,threeparttable, booktabs,mathtools,dsfont}

@@

% syntax and readability
\newcommand{\para}{\bigskip\noindent}
\renewcommand{\(}{\left(}
\renewcommand{\)}{\right)}
\renewcommand{\[}{\left[}
\renewcommand{\]}{\right]}
\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}} 
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}

% settings
\graphicspath{{/Users/magnusmunch/Documents/OneDrive/PhD/EBEN/graphs/}}
\setlength{\evensidemargin}{.5cm} \setlength{\oddsidemargin}{.5cm}
\setlength{\textwidth}{15cm}
\setlength{\parindent}{0pt}
\title{Adaptive group-regularized logistic elastic net regression Supplementary material}
\date{}
\author{Magnus M. M\"unch$^{1,2}$\footnote{Correspondence to: \href{mailto:m.munch@vumc.nl}{m.munch@vumc.nl}}, Carel F.W. Peeters$^{1}$, Aad W. van der Vaart$^{2}$ \\ and Mark A. van de Wiel$^{1,3}$}

\begin{document}
\maketitle
\noindent
1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health research institute, VU University Medical Center, PO Box 7057, 1007 MB
Amsterdam, The Netherlands\\
2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\
3. Department of Mathematics, VU University, Amsterdam, The Netherlands

\para
\textbf{Keywords}: Empirical Bayes; High-dimensional data; Prediction; Variational Bayes \\

\para
\textbf{Software available from}: \url{https://github.com/magnusmunch/gren/}\\

\section{Introduction}
This document contains supplementary material to the paper `Adaptive group-regularized logistic elastic net regression'. Section \ref{sec:enetprior} contains extra details on the elastic net prior distribution. Section \ref{sec:objectivefunction} gives the objective function that is maximised in the variational Bayes procedure introduced in Section 3 of the Main Document. Section \ref{sec:polyagamma} gives details on the P{\'o}lya-Gamma parametrisation introduced in Section 3 of the Main Document. The derivations of the variational posterior used in the Main Document are shown in Section \ref{sec:variationaldistributions}. Section \ref{sec:computationalcomplexity} shows how we may reduce the computational complexity of our variational Bayes procedure. Section \ref{sec:expectedloglikelihood} elaborates on the Variational Bayes expected joint log likelihood that is maximised to estimate the penalty multipliers as explained in Section 3 of the Main Document. The full procedure of the proposed method is shown in Section \ref{sec:fullestimationprocedure}. We give some extra results on the microRNA in colorectal cancer data application in the Main Document in Section \ref{sec:extra} \cite[]{neerincx_combination_2018}. Finally, we introduce the ridge and lasso versions of the proposed method in Section \ref{sec:evbridgelasso}.

\section{The elastic net prior}\label{sec:enetprior}
\subsection{Density function}
The density function of $\beta$, drawn from an elastic net prior distribution, is given by:

$$
f(\beta) = g(\lambda_1,\lambda_2) \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\].
$$

The normalizing constant here is calculated as:

$$
g(\lambda_1,\lambda_2) =\left\{ \int_{-\infty}^{\infty} \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta \right\}^{-1}= \frac{\sqrt{\lambda_2}}{2} \phi \( \frac{\lambda_1}{2 \sqrt{\lambda_2}} \) \Phi \(\frac{-\lambda_1}{2 \sqrt{\lambda_2}} \)^{-1},
$$

with $\phi(x)$ and $\Phi(x)$ the standard normal density and distribution functions of $x$, respectively.

\subsection{Expectation and variance}
The expectation and variance of $\beta$ are:

\begin{align*}
\E[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta \cdot \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta=0, \\
\V[\beta] &= g(\lambda_1,\lambda_2) \int_{-\infty}^{\infty} \beta^2 \cdot \exp\[-\frac{1}{2} (\lambda_1 |\beta| + \lambda_2 \beta^2)\] d\beta - \E[\beta]^2\\ &= \frac{\lambda_1^2}{4 \lambda_2^2} + \frac{1}{\lambda_2} - \frac{\lambda_1}{2 \lambda_2^{3/2}} \phi \( \frac{\lambda_1}{2 \sqrt{\lambda_2}} \) \Phi \(\frac{-\lambda_1}{2 \sqrt{\lambda_2}} \)^{-1}.
\end{align*}

\subsection{Simulation}
Generating samples from the elastic net prior is not straightforward. \cite{li_bayesian_2010} show that the elastic net prior may be written as a scale mixture of normals, with mixing parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} \tr$:

\begin{align*}
\beta | \tau & \sim \mathcal{N} \left(0,\frac{1}{\lambda_2} \frac{\tau - 1}{\tau} \right), \\
\tau & \sim \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2}{\lambda_1^2}, \left(1,\infty \right) \right),
\end{align*}

where $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$. So with an efficient truncated gamma sampler available we may sample from the prior using any standard software package that includes a Gaussian sampling routine. More specifically, we require a a sample from the $\mathcal{TG} \left( \frac{1}{2},8 \lambda2/\lambda_1^2, \left(1,\infty \right) \right)$ distribution. The CDF of this distribution is given by:

$$
F_{T}(\tau) =	
\begin{dcases}
0, & \text{if }\tau < 1 \\
1 - \frac{\Phi\(-\frac{\lambda_1}{2} \sqrt{\frac{\tau}{\lambda_2}} \)}{\Phi\(-\frac{\lambda_1}{2 \sqrt{\lambda_2}} \)}, & \text{if } \tau \geq 1.
\end{dcases}
$$

Since $F_{T}(\tau)$ is continuous and strictly monotonically increasing for $\tau \geq 1$, the quantile function may be written as a function of the probability $p \in (0,1)$:

$$
Q_{T}(p) =	F^{-1}_{T}(p) = \frac{4 \lambda_2}{\lambda_1^2} \Phi^{-1} \[ (1 - p) \Phi\(-\frac{\lambda_1}{2 \sqrt{\lambda_2}} \)\]^2,
$$

where $\Phi^{-1}(p)$ is the quantile function of the standard normal distribution, i.e, the probit function. Employing inverse transform sampling for the truncated gamma distribution, an elastic net prior sampling scheme is now: \\

\begin{algorithmic}
	\State{Generate $U \sim \mathcal{U} (0,1)$}
	\State{Set $\tau = \frac{4 \lambda_2}{\lambda_1^2} \Phi^{-1} \[ (1 - u) \Phi\(-\frac{\lambda_1}{2 \sqrt{\lambda_2}} \)\]^2$}
	\State{Set $\nu^2= \frac{\tau - 1}{\tau \lambda_2}$}
	\State{Generate $\beta \sim \mathcal{N}(0,\nu^2)$}
	\State{\Return{$\beta$}}
\end{algorithmic}

\section{Objective function}\label{sec:objectivefunction}

The lower bound on the marginal likelihood, or evidence lower bound (ELBO), at iteration $t$ is given by:

\begin{subequations}
	\begin{align}
	\text{ELBO} (Q^{(t)}) & = \E_{Q^{(t)}} [\log \L (\y, \bomega, \bbeta, \btau)] - \E_{Q^{(t)}} [\log Q (\bomega, \bbeta, \btau)] \\
	& = \E_{Q^{(t)}} [\log \L (\y ; \bbeta)] + \E_{Q^{(t)}} [\log \pi(\bomega | \bbeta)] - \E_{Q^{(t)}} [\log q(\bomega)] \label{eq:elboomega} \\
	& \,\,\,\,\, + \E_{Q^{(t)}} [\log \pi(\bbeta|\btau)] - \E_{Q^{(t)}} [\log q(\bbeta)] \label{eq:elbobeta}\\
	& \,\,\,\,\, + \E_{Q^{(t)}} [\log \pi(\btau)] - \E_{Q^{(t)}} [\log q(\btau)] \label{eq:elbotau}.
	\end{align}
\end{subequations}

We start with the computation of (\ref{eq:elboomega}). To this end we define:

$$
h(m_i, \omega_i) := \sum_{k=0}^{\infty} (-1)^k \frac{\Gamma(k+m_i}{k + 1}\frac{2k +m_i}{\sqrt{2 \pi \omega_i^3}}\exp\[-\frac{(2k + m_i)^2}{8 \omega_i} \].
$$

Let $\c = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix} \tr$. Furthermore, in the following, all expectations and variances are with respect to $Q^{(t)}$. Next we write out the three terms in (\ref{eq:elboomega}):

\begin{subequations} \label{eq:elboomega2}
	\begin{align}
	\E [\log \L (\y ; \bbeta)] & = \sum_{i=1}^n \log \binom{m_i}{y_i} + \y \tr \X \E [\bbeta] - \m \tr \E \left\{ \log [1 + \exp(\x_i \tr \bbeta)] \right\}, \\
	\E [\log \pi(\bomega | \bbeta)] & = \m \tr \E \left\{ \log [1 + \exp(\x_i \tr \bbeta)] \right\} - n \log 2 - \frac{1}{2} \m \tr \X \E [\bbeta] \\
	& \,\,\,\,\, - \sum_{i=1}^n \log \Gamma (m_i) - \frac{1}{2} \sum_{i=1}^n [ ( \x_i \tr \E [\bbeta] )^2 + \x_i \tr \V [\bbeta] \x_i \tr] \E [\omega_i] \\
	& \,\,\,\,\,+ \sum_{i=1}^n \E \left\{ \log [ h(m_i, \omega_i) ] \right\}, \\
	\E [\log q(\bomega)] & =  \m \tr \log [1 + \exp(\c)] - n \log 2 - \frac{1}{2} \m \tr \c - \sum_{i=1}^n \log \Gamma (m_i) - \frac{1}{2} \sum_{i=1}^n c_i^2 \E [\omega_i] \\
	& \,\,\,\,\, + \sum_{i=1}^n \E \left\{ \log [ h(m_i, \omega_i) ] \right\},
	\end{align}
\end{subequations}

where all mathematical operations on vectors and matrices are element wise. After combining the terms in (\ref{eq:elboomega2}) and substituting the expectations and variances given in the Main Document, several terms cancel to give:

\begin{subequations} \label{eq:elboomega3}
	\begin{align}
	\E_{Q^{(t)}} & [\log \L (\y ; \bbeta)] + \E_{Q^{(t)}} [\log \pi(\bomega | \bbeta)] - \E_{Q^{(t)}} [\log q(\bomega)] \\ & = \sum_{i=1}^n \log \binom{m_i}{y_i} + ( \y - \frac{1}{2} \m) \tr \X \bmu^{(t)} + \m \tr \left\{ \frac{1}{2} \c^{(t)} - \log [1 + \exp(\c^{(t)})] \right\} \label{eq:elboomega4}\\
	& \,\,\,\,\, + \frac{1}{4} \sum_{i=1}^n \frac{m_i}{c_i^{(t)}} \tanh \(\frac{c_i^{(t)}}{2} \)\left\{ (c_i^{(t)})^2 - ( \x_i \tr \bmu^{(t)} )^2 - \x_i \tr \bSigma^{(t)} \x_i \tr \right\}.
	\end{align}
\end{subequations}

Inspection of the updating equations in the Main Document, learns us that the last term in the right-hand side of (\ref{eq:elboomega3}) equals zero, so that we are left with just (\ref{eq:elboomega4}). After a change of variables $\psi_j = \tau_j - 1$, the two terms in (\ref{eq:elbobeta}) are as follows:

\begin{subequations} \label{eq:elbobeta2}
	\begin{align}
	\E [\log \pi (\bbeta | \btau)] & = \frac{1}{2}\sum_{g=1}^G |\mathcal{G}(g)| \log \lambda'_g + \frac{p}{2} \log \lambda_2 - \frac{p}{2} \log (2 \pi) \\
	& \,\,\,\,\, - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j}}\) (\bmu_j^2 + \bSigma_{jj}) \\
	& \,\,\,\,\, - \frac{1}{2} \sum_{j=1}^p \E [\log \psi_j] + \frac{1}{2} \sum_{j=1}^p \E [ \log(\psi_j + 1)], \\
	\E [\log q (\bbeta )] & = -\frac{p}{2} - \frac{p}{2}\log (2 \pi) - \frac{p}{2} \log |\bSigma| .
	\end{align}
\end{subequations}

We combine the terms in (\ref{eq:elbobeta2}) and arrive at the following for (\ref{eq:elbobeta}):

\begin{subequations} \label{eq:elbobeta3}
	\begin{align}
	\E_{Q^{(t)}} & [\log \pi (\bbeta | \btau)] - \E_{Q^{(t)}} [\log q (\bbeta )] \\ 
	& = \frac{1}{2}\sum_{g=1}^G |\mathcal{G}(g)| \log \lambda'^{(t)}_g + \frac{p}{2} \log \lambda_2 + \frac{p}{2} \log |\bSigma^{(t)}| + \frac{p}{2} \\
	& \,\,\,\,\, - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'^{(t)}_g \sum_{j \in \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j^{(t)} } }\) \((\bmu^{(t)}_j)^2 + \bSigma_{jj}^{(t)}\) \\
	& \,\,\,\,\, - \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log \psi_j] + \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [ \log(\psi_j + 1)].
	\end{align}
\end{subequations}

Repeating the same exercise for (\ref{eq:elbotau}) gives:

\begin{subequations} \label{eq:elbotau2}
	\begin{align}
	\E [\log \pi (\btau)] & = p \log \lambda_1 - \frac{p}{2} \log \lambda_2 - \frac{5 p }{2} \log 2 - \frac{p}{2} \log \pi - \frac{p}{2} - \frac{p \lambda_1^2}{8 \lambda_2} \\
	& \,\,\,\,\, - p \log \Phi \( \frac{-\lambda_1}{2 \sqrt{ \lambda_2}} \) - \frac{\lambda_1}{4 \sqrt{\lambda_2}} \sum_{j=1}^p \chi_j^{1/2} - \frac{1}{2} \sum_{j=1}^p \E [\log ( \psi_j + 1)], \\
	\E [\log q (\btau)] & = p \log \lambda_1 - \frac{p}{2} \log \lambda_2 - \frac{3p}{2} \log 2 -\frac{p}{2} \log \pi -\frac{p}{2} - \frac{\lambda_1}{2 \sqrt{\lambda_2}} \sum_{j=1}^p \chi_j^{1/2} \\
	& \,\,\,\,\, + \frac{1}{4} \sum_{j=1}^p \log \chi_j - \frac{1}{2} \sum_{j=1}^p \E [\log \psi_j].
	\end{align}
\end{subequations}

Again, we combine the terms in (\ref{eq:elbotau2}) to get:

\begin{subequations} \label{eq:elbotau3}
	\begin{align}
	\E_{Q^{(t)}} & [\log \pi (\btau)] - \E_{Q^{(t)}} [\log q (\btau)] 
	\\ & = p \log 2 - \frac{p \lambda_1^2}{8 \lambda_2} - p \log \Phi \( \frac{-\lambda_1}{2 \sqrt{ \lambda_2}} \) + \frac{\lambda_1}{4 \sqrt{\lambda_2}} \sum_{j=1}^p \sqrt{\chi_j^{(t)}} - \frac{1}{4} \sum_{j=1}^p \log \chi_j^{(t)} \\
	& \,\,\,\,\, + \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log \psi_j] - \frac{1}{2} \sum_{j=1}^p \E_{Q^{(t)}} [\log ( \psi_j + 1)].
	\end{align}
\end{subequations}

Adding up the terms in (\ref{eq:elboomega3}), (\ref{eq:elbobeta3}), and (\ref{eq:elbotau3}) gives the ELBO, at iteration $t$:

\begin{align*}
\text{ELBO} (Q^{(t)}) & \propto (\y - \frac{1}{2} \m ) \tr \X \bmu^{(t)} + \m \tr \{ \frac{1}{2} \c^{(t)} - \log [ \exp(\c^{(t)}) + 1 ]\} + \frac{1}{2} \sum_{g=1}^G |\G(g)| \log \lambda'^{(t)}_g \\ 
& + \frac{1}{2}\log |\bSigma^{(t)}| - \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'^{(t)}_g \sum_{j \in \mathcal{G}(g)} \(1 + \frac{\lambda_1}{2 \sqrt{\lambda_2 \chi_j^{(t)} } }\) \((\bmu^{(t)}_j)^2 + \bSigma_{jj}^{(t)}\) \\
& + \frac{\lambda_1}{4 \sqrt{\lambda_2}} \sum_{j=1}^p \sqrt{\chi_j^{(t)}} - \frac{1}{4} \sum_{j=1}^p \log \chi_j^{(t)}.
\end{align*}

Here, proportionality is with respect to the variational and penalty parameters.
